---
title: "EEMS_Calcar"
author: "Jon Puritz"
date: "1/15/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Estimating effective migration surfaces [(EEMS)](https://github.com/dipetkov/eems)

Here is the code used to run eems_sats in command-line and plot the results in R. Detailed steps for running this program can be accessed [here](https://github.com/dipetkov/eems/tree/master/runeems_sats).

The program runeems_sats implements the EEMS method for analyzing spatial population structure. This version uses raw microsatellite data.

### Input data format (microsatellites)
`runeems_sats` requires three data input files that have the same file name but different extension. For this dataset, the file name is `microsat` + extension.

#### microsat.sites
`microsat.sites` is the matrix of allele copies; missing alleles are specified by any negative number. For n individuals at L loci, the sites matrix is n-by-L for L haploid markers and n-by-2L for L diploid markers. The matrix of allele copies was generated from `Genepop.txt`.

#### microsat.coord
`microsat.coord` are the sample coordinates, two coordinates per sample, one sample per line. The sampling locations should be given in the same order as the rows and columns of the dissimilarity matrix. Coordinates were taken from `Calcar_TESS2 - randGeo.txt`.

#### microsat.outer
`microsat.outer` are the habitat coordinates, as a sequence of vertices that form a closed polygon. The habitat vertices should be listed counterclockwise and the first vertex should also be the last vertex, so that the outline is a closed ring. Otherwise, EEMS attempts to "correct" the polygon and prints a warning message. Habitat coordinates were made by A. Zyck based on site coordinates.

### All files are included in the repository

## Run Parameters

```{bash}
cat params-microsat-D48-chain1.ini
```

```{bash eval=FALSE}
cp params-microsat-D48-chain1.ini params-microsat-D48-chain2.ini
cp params-microsat-D48-chain1.ini params-microsat-D48-chain3.ini
sed -i 's/chain1/chain2/g' params-microsat-D48-chain2.ini
sed -i 's/chain1/chain3/g' params-microsat-D48-chain3.ini
```

```{bash eval=FALSE}
for i in {96,192,384,768}
do
  for j in {1,2,3}
  do
  cp params-microsat-D48-chain1.ini params-microsat-D$i-chain$j.ini
  sed -i "s/chain1/chain$j/g" params-microsat-D$i-chain$j.ini
  sed -i "s/D48/D$i/g" params-microsat-D$i-chain$j.ini
  sed -i "s/nDemes = 48/nDemes = $i/g" params-microsat-D$i-chain$j.ini
  done
done
```

## Run eems in parallel 3 chains per deme level
```{bash eval=FALSE}
ls params-microsat-D* | parallel runeems_sats --params {} >eems.out 2>&1
```



```{r}
library("rEEMSplots")
library("ggplot2")
library("dplyr")
library("rworldmap")
library("rworldxtra")
```


```{r}
path = "/home/jpuritz/eems_calcar/"
plotpath = "/home/jpuritz/eems_calcar/eems"
projection_none <- "+proj=longlat +datum=WGS84"
projection_mercator <- "+proj=merc +datum=WGS84"
```

```{r}
dirs = c(paste0(path,"microsat-D48-chain1"),paste0(path,"microsat-D48-chain2"),paste0(path,"microsat-D48-chain3"),paste0(path,"microsat-D96-chain1"),paste0(path,"microsat-D96-chain2"),paste0(path,"microsat-D96-chain3"),paste0(path,"microsat-D192-chain1"),paste0(path,"microsat-D192-chain2"),paste0(path,"microsat-D192-chain3"),paste0(path,"microsat-D384-chain1"),paste0(path,"microsat-D384-chain2"),paste0(path,"microsat-D384-chain3"),paste0(path,"microsat-D768-chain1"),paste0(path,"microsat-D768-chain2"),paste0(path,"microsat-D768-chain3"))

coord <- read.table(paste0(path, "microsat.coord"))
coord.dec <- SpatialPoints(cbind(coord$V2, coord$V1), proj4string=CRS(projection_none))
coord.merc <- spTransform(coord.dec, CRS(projection_mercator))

eems.plots(mcmcpath = dirs, plotpath,longlat = FALSE,add.grid=F,add.outline = T,add.demes = F, projection.in = projection_none, projection.out = projection_mercator, add.map = T,add.abline = T, add.r.squared = T,add.title = FALSE,m.plot.xy = { points(coord.merc, col = "black", cex =1, pch=19) },q.plot.xy = { points(coord.merc, col = "black", cex =1, pch=19) })
```


### EEMS runs per deme setting

#### 48

```{r}
dirs = c(paste0(path,"microsat-D48-chain1"),paste0(path,"microsat-D48-chain2"),paste0(path,"microsat-D48-chain3"))
plotpath = "/home/jpuritz/eems_calcar/eems-D48"

eems.plots(mcmcpath = dirs, plotpath,longlat = FALSE, add.grid=F,add.outline = T,add.demes = F, projection.in = projection_none, projection.out = projection_mercator, add.map = T,add.abline = T, add.r.squared = T,add.title = FALSE,m.plot.xy = { points(coord.merc, col = "black", cex =1, pch=19) },q.plot.xy = { points(coord.merc, col = "black", cex =1, pch=19) })
```

#### 96

```{r}
dirs = c(paste0(path,"microsat-D96-chain1"),paste0(path,"microsat-D96-chain2"),paste0(path,"microsat-D96-chain3"))
plotpath = "/home/jpuritz/eems_calcar/eems-D96"

eems.plots(mcmcpath = dirs, plotpath,longlat = FALSE, add.grid=F,add.outline = T,add.demes = F, projection.in = projection_none, projection.out = projection_mercator, add.map = T,add.abline = T, add.r.squared = T,add.title = FALSE,m.plot.xy = { points(coord.merc, col = "black", cex =1, pch=19) },q.plot.xy = { points(coord.merc, col = "black", cex =1, pch=19) })

```

#### 192

```{r}
dirs = c(paste0(path,"microsat-D192-chain1"),paste0(path,"microsat-D192-chain2"),paste0(path,"microsat-D192-chain3"))
plotpath = "/home/jpuritz/eems_calcar/eems-D192"

eems.plots(mcmcpath = dirs, plotpath,longlat = FALSE, add.grid=F,add.outline = T,add.demes = F, projection.in = projection_none, projection.out = projection_mercator, add.map = T,add.abline = T, add.r.squared = T,add.title = FALSE,m.plot.xy = { points(coord.merc, col = "black", cex =1, pch=19) },q.plot.xy = { points(coord.merc, col = "black", cex =1, pch=19) })

```

#### 384

```{r}
dirs = c(paste0(path,"microsat-D384-chain1"),paste0(path,"microsat-D384-chain2"),paste0(path,"microsat-D384-chain3"))
plotpath = "/home/jpuritz/eems_calcar/eems-D384"

eems.plots(mcmcpath = dirs, plotpath,longlat = FALSE, add.grid=F,add.outline = T,add.demes = F, projection.in = projection_none, projection.out = projection_mercator, add.map = T,add.abline = T, add.r.squared = T,add.title = FALSE,m.plot.xy = { points(coord.merc, col = "black", cex =1, pch=19) },q.plot.xy = { points(coord.merc, col = "black", cex =1, pch=19) })

```


#### 768

```{r}
dirs = c(paste0(path,"microsat-D768-chain1"),paste0(path,"microsat-D768-chain2"),paste0(path,"microsat-D768-chain3"))
plotpath = "/home/jpuritz/eems_calcar/eems-D768"

eems.plots(mcmcpath = dirs, plotpath,longlat = FALSE, add.grid=F,add.outline = T,add.demes = F, projection.in = projection_none, projection.out = projection_mercator, add.map = T,add.abline = T, add.r.squared = T,add.title = FALSE,m.plot.xy = { points(coord.merc, col = "black", cex =1, pch=19) },q.plot.xy = { points(coord.merc, col = "black", cex =1, pch=19) })

```


